Solving Collective Spin Hamiltonians with Decoherence

### 2018-08-20 soc_squeezing/cumulant_expansions.tex
changed strategy for simplifying last commutator in Appendix A

sums to get that for $p>1$

@@

  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}

@@

  \label{eq:E_mn_p_semifinal}

@@

e.g.~$\Z_{\ell+1}=\set{0,1,\cdots,\ell}$.  We now expand

@@

  \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z}
  = \sum_{q=0}^{p-1} \p{2\mu S_z}^q
  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}} \prod_{j\in c} \p{2k_{j+1}+j+1-n},
  \label{eq:product_expansion}

@@

where $\C\p{k,\S}$ denotes the set of $k$-combinations (i.e.~subsets
with $k$ elements) of $\S$; the number of elements in
$\C_{p-1-q}\p{\Z_{p-1}}$ is
${ p-1 \choose p-1-q } = { p-1 \choose q }$.  Substituting
\eqref{eq:D_mn_k} and \eqref{eq:product_expansion} into
\eqref{eq:E_mn_p_semifinal}, we get that for $p>1$,

@@

  = n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1}
  \sum_{s=0}^2 f_{mnp\ell}^{(s)} \p{2\mu S_z}^s
  \sum_{q=0}^{p-1} g_{np\ell q} \p{2\mu S_z}^q,
  \label{eq:E_mn_p}

@@

where

@@

  f_{mnp\ell}^{(0)}
  &= \p{m-p-\ell} \p{n-p} \p{\ell+m-n} \p{2\ell-n+p}, \\
  f_{mnp\ell}^{(1)}
  &= \p{m-p-\ell} \p{n-p} \p{3\ell+m-2n+p}, \\
  f_{mnp\ell}^{(2)}
  &= \p{m-p-\ell} \p{n-p}, \\
  g_{np\ell q}
  &=
  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}}
  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}
  \prod_{j\in c} \p{2k_{j+1}+j+1-n},

@@

and for $p=1$ we can resolve \eqref{eq:E_mn_1_first} to get

@@

  = \f16 mn \p{m-1} \p{3n-m-1} S_\bmu^{n-1} S_\mu ^{m-1} \p{m+2\mu S_z}.
  \label{eq:E_mn_1}

@@

With reference to \eqref{eq:E_mn_p}-\eqref{eq:E_mn_1}, we finally have

@@

  = mn S_\bmu^{n-1} S_\mu^{m-1}
  \sum_{q\in\set{0,1}} \p{m-n}^{1-q}\p{2\mu S_z}^q
  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)}.

@@

>>>>

sums to get that

@@

  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le\cdots\le k_{p-1}}}

@@

@@

e.g.~$\Z_{\ell+1}=\set{0,1,\cdots,\ell}$.  Defining for shorthand

@@

  \sum_{\p{\v k,p,\ell}} X \equiv
  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le\cdots\le k_{p-1}}} X,
  F_{\mu np} \equiv p - n + 2\mu S_z,

@@

we can say

@@

  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+F_{\mu np}}
  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}.
  \label{eq:E_mn_p_k_sum}

@@

We now define

@@

  f_{\mu np\ell}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{\mu np}},

@@

and simplify

@@

  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  &= \sum_{\p{\v k;p-1;\ell}}
  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  f_{\mu n\ell}\p{k_{p-2},1} \\
  &= \f12 \sum_{\p{\v k,p-2,\ell}}
  \prod_{j=3}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  f_{\mu np\ell}\p{k_{p-3},1} f_{\mu np\ell}\p{k_{p-3},2} \\
  &\qquad \vdots \tag*{} \\
  &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
  \prod_{j=1+r}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
  \prod_{q=1}^r f_{\mu np\ell}\p{k_{p-1-r},q} \\
  &\qquad \vdots \tag*{} \\
  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{\mu np\ell}\p{0,q} \\
  &= { \ell + p - 1 \choose p - 1 }
  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}}.

@@

Substitution into \eqref{eq:E_mn_p_k_sum} yields

@@

  = n^{\underline{p}} \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)}
  { \ell + p - 1 \choose p - 1 }
  \prod_{q=0}^{p-1} \p{\ell-q+F_{\mu np}}
  = n^{\underline{p}} S_\bmu^{n-1} S_\mu^{m-1} G_{\mu mnp}

@@

for
  G_{\mu mnp}
  \equiv \sum_{\ell=0}^{m-p-1} \p{m-p-\ell} \p{n-p}
  \p{\ell+m-p+F_{\mu np}}
  { \ell + p - 1 \choose p - 1 }
  \prod_{q=0}^{p-1} \p{\ell-q+F_{\mu np}}.

\begin{align}
  \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z}
  = \sum_{q=0}^{p-1} \p{2\mu S_z}^q
  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}} \prod_{j\in c} \p{2k_{j+1}+j+1-n},
  \label{eq:product_expansion}
\end{align}
where $\C\p{k,\S}$ denotes the set of $k$-combinations (i.e.~subsets
with $k$ elements) of $\S$; the number of elements in
$\C_{p-1-q}\p{\Z_{p-1}}$ is
${ p-1 \choose p-1-q } = { p-1 \choose q }$.  Substituting
\eqref{eq:D_mn_k} and \eqref{eq:product_expansion} into
\eqref{eq:E_mn_p_semifinal}, we get that for $p>1$,
\begin{align}
  E_{\mu mn}^{(p)}
  = n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1}
  \sum_{s=0}^2 f_{mnp\ell}^{(s)} \p{2\mu S_z}^s
  \sum_{q=0}^{p-1} g_{np\ell q} \p{2\mu S_z}^q,
  \label{eq:E_mn_p}
\end{align}
where
\begin{align}
  f_{mnp\ell}^{(0)}
  &= \p{m-p-\ell} \p{n-p} \p{\ell+m-n} \p{2\ell-n+p}, \\
  f_{mnp\ell}^{(1)}
  &= \p{m-p-\ell} \p{n-p} \p{3\ell+m-2n+p}, \\
  f_{mnp\ell}^{(2)}
  &= \p{m-p-\ell} \p{n-p}, \\
  g_{np\ell q}
  &=
  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}}
  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}
  \prod_{j\in c} \p{2k_{j+1}+j+1-n},
\end{align}
and for $p=1$ we can resolve \eqref{eq:E_mn_1_first} to get
\begin{align}
  E_{\mu mn}^{(1)}
  = \f16 mn \p{m-1} \p{3n-m-1} S_\bmu^{n-1} S_\mu ^{m-1} \p{m+2\mu S_z}.
  \label{eq:E_mn_1}
\end{align}
With reference to \eqref{eq:E_mn_p}-\eqref{eq:E_mn_1}, we finally have
\begin{align}
  \sp{S_\mu^m, S_\bmu^n}_-
  = C_{\mu mn}
  = mn S_\bmu^{n-1} S_\mu^{m-1}
  \sum_{q\in\set{0,1}} \p{m-n}^{1-q}\p{2\mu S_z}^q
  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)}.
\end{align}
